---
title: |
  | **Ordinary Differential Equation**
  | **Tentative Topics Schedule**
author: |
  | MTH-321 Sections B & C
  | Spring 2024
  | University of Portland
bibliography: references.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(file.path("R","functions.R"))
```

```{r echo=FALSE}
create_footer()
```

<br>

> See Books & Online Resources Lists for the readings & practice materials.
>
> The readings & practice materials are not mandatory but it is encouraged.

<br>The "Reading & Practice" column in the table below contains page numbers on which it refers to a label in the Books & Online Resources List. For example "Pgs. 1-5 [H]" refers to pages 1-5 of the first item in the list, which is the textbook titled "Ordinary Differential Equations: A Primer on Dynamics and Systems".

<br>Most of the materials (worksheets and homework) of each topic was taken from Inquiry oriented differential equations (IODE) by @iode2018.

<br>

## **Topics and Materials**

| **Week** | **Day** | **Topic** | **Activity** | **Homework** | **Reading & Practice** |
|:---:|:---:|:---:|:---:|:---:|:---:|
| **1** | Tu 1/16 | Introduction and Orientation to ODES,<br>Classification of Equations, &<br>Principles of Solutions | Algebra & Calculus Review | Assigned: Homework 0 | Syllabus |
|  | Th 1/18 | 1st-Order ODEs,<br>Modeling Physical Processes, &<br>Existence and Uniqueness | TBA | - | TBA |
| **2** | Tu 1/23 | Analyzing Equilibriums to Linear and Non-Linear 1st-Order ODEs | TBA | Assigned: Homework 1 | TBA |
|  | Th 1/25 | Bifurcations in One Dimension | TBA | - | TBA |
| **3** | Tu 1/30 | Separation of Variables for Solving 1st-Order ODEs | TBA | Assigned: Homework 2 | TBA |
|  | Th 2/1 | Integrating Factors for Solving 1st-Order ODEs | TBA | - | TBA |
| **4** | Tu 2/6 | Laplace Transforms | TBA | Assigned: Homework 3 | TBA |
|  | Th 2/8 | Laplace Transforms for Solving 1st-Order ODEs | TBA | - | TBA |
| **5** | Tu 2/13 | _Review_ | Exam 1 Examples | - | Exam 1 Guide |
|  | Th 2/15 | **Exam 1** | - | - | - |
| **6** | Tu 2/20 | Systems of 1st-Order ODEs,<br>Modeling Physical Processes, &<br>Existences and Uniqueness | TBA | Assigned: Homework 4 | TBA |
|  | Th 2/22 | Analyzing Equilibriums to 1st-Order Linear and Non-Linear Systems | TBA | - | TBA |
| **7** | Tu 2/27 | Linear Transformations & Eigentheory | TBA | Assigned: Homework 5 | TBA |
|  | Th 2/29 | Stability Analysis of 1st-Order Linear Systems | TBA | - | TBA |
| **8** | Tu 3/5 | _Spring Vacation_ | - | - | - |
|  | Th 3/7 | _Spring Vacation_ | - | - | - |
| **9** | Tu 3/12 | Bifurcations in Two Dimensions | TBA | Assigned: Homework 6 | TBA |
|  | Th 3/14 | Eigenvalues and Eigenvectors for solving 1st-Order Linear Systems | TBA | - | TBA |
| **10** | Tu 3/19 | _Review_ | Exam 2 Examples | - | Exam 2 Guide |
|  | Th 3/21 | **Exam 2** | - | - | - |
| **11** | Tu 3/26 | Higher Order ODEs,<br>Modeling Physical Processes, &<br>Characteristic Equations | TBA | Assigned: Homework 7 | TBA |
|  | Th 3/28 | Analyzing Linear and Non-Linear 2nd-Order ODEs | TBA | - | TBA |
| **12** | Tu 4/2 | Undetermined Coefficients for Solving Linear 2nd-Order ODEs | TBA | Assigned: Homework 8 | TBA |
|  | Th 4/4 | Variation of Parameters for Solving Linear 2nd-Order ODEs | TBA | - | TBA |
| **13** | Tu 4/9 | Laplace Transforms for Solving Linear 2nd-Order ODEs | TBA | Assigned: Homework 9 | TBA |
|  | Th 4/11 | Power and Fourier Series Methods for Solving Linear 2nd-Order ODEs | TBA | - | TBA |
| **14** | Tu 4/16 | _Review_ | Exam 3 Examples | - | Exam 3 Guide |
|  | Th 4/18 | **Exam 3** | - | - | - |
| **15** | Tu 4/23 | Euler's Method for Solving ODEs | TBA | Assigned: Homework 10 | TBA |
|  | Th 4/25 | Runge-Kutta Method for Solving ODEs | TBA | - | TBA |
| **16** | Tu 4/30 | **Final Exam Section B** | - | - | - |
|  | Th 5/2 | **Final Exam Section C** | - | - | - |

<br>

## **Books & Online Resources Lists**

```{r echo=FALSE}
refs <- bibtex::read.bib("references.bib")
```

*Click on the link to access the resources.*

### Textbooks

[H] [`r capture.output(print(refs["hallstrom2018"]))`](https://drive.google.com/file/d/12G9cs0RQPqiXb7vjJXVH58ePJ8DZBgC1/view?usp=sharing){target="_blank"}

[T] [`r capture.output(print(refs["trench2013elementary"]))`](https://digitalcommons.trinity.edu/cgi/viewcontent.cgi?article=1007&context=mono){target="_blank"}

### Websites

[C] `r capture.output(print(refs["clontz2022"]))`

[P] `r capture.output(print(refs["paulodenotes"]))`

<br>

# References

<div id="refs"></div>

<br>
