---
title: |
  | **Ordinary Differential Equation**
  | **Tentative Topics Schedule**
author: |
  | MTH 321 Sections B & C
  | Spring 2024
  | University of Portland
bibliography: references.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(file.path("R","functions.R"))
```

```{r echo=FALSE}
create_footer()
```

<br>

> See Books & Online Resources Lists for the readings & practice materials.
> 
> The reading materials are not mandatory but it is encouraged.

<br>The "Reading" column in the table below contains page numbers on which it refers to a label in the Books & Online Resources List. For example "Pgs. 1-5 [H]" refers to pages 1-5 of the first item in the list, which is the textbook titled "Ordinary Differential Equations: A Primer on Dynamics and Systems".

<br>

## **Topics and Materials**

| **Week** | **Day** | **Topic** | **Worksheet** | **Homework** | **Reading** |
|:-----:|:-----:|:---------------:|:-----:|:-----:|:-----:|
| **1** | Tu 1/16 | _Cancelled_ | - | - | - |
|  | Th 1/18 | Introduction and Orientation to ODES, <br>Classification of Equations, & <br>Principles of Solutions | Algebra & Calculus Review<br>Verifying Solutions | - | TBA |
| **2** | Tu 1/23 | 1st-Order ODEs, <br>Existence and Uniqueness, <br>Modeling Physical Systems, & <br>Analyzing Equilibriums of Linear and Non-Linear 1st-Order ODEs | TBA | Assigned: Homework 1 | TBA |
|  | Th 1/25 | Separation of Variables & <br>Integrating Factors for Solving 1st-Order ODEs | TBA | - | TBA |
| **3** | Tu 1/30 | Laplace Transforms | TBA | Assigned: Homework 2 | TBA |
|  | Th 2/1 | Laplace Transforms for Solving 1st-Order ODEs | TBA | - | TBA |
| **4** | Tu 2/6 | Euler's Method for solving ODEs | TBA | Assigned: Homework 3 | TBA |
|  | Th 2/8 | Bifurcations in One Dimension | TBA | - | TBA |
| **5** | Tu 2/13 | _Review_ | Exam 1 Examples | - | Exam 1 Guide |
|  | Th 2/15 | **Exam 1** | - | - | - |
| **6** | Tu 2/20 | Systems of 1st-Order ODEs,  <br>Existences and Uniqueness, & <br>Modeling Physical Processes | TBA | Assigned: Homework 4 | TBA |
|  | Th 2/22 | Analyzing Equilibriums to 1st-Order Linear and Non-Linear Systems | TBA | - | TBA |
| **7** | Tu 2/27 | Linear Transformations & <br>Eigenthoery | TBA | Assigned: Homework 5 | TBA |
|  | Th 2/29 | Stability Analysis of 1st-Order Linear Systems | TBA | - | TBA |
| **8** | Tu 3/5 | _Spring Vacation_ | - | - | - |
|  | Th 3/7 | _Spring Vacation_ | - | - | - |
| **9** | Tu 3/12 | Eigenvalues and Eigenvectors for solving 1st-Order Linear Systems | TBA | Assigned: Homework 6 | TBA |
|  | Th 3/14 | Bifurcations in Two Dimensions | TBA | - | TBA |
| **10** | Tu 3/19 | _Review_ | Exam 2 Examples | - | Exam 2 Guide |
|  | Th 3/21 | **Exam 2** | - | - | - |
| **11** | Tu 3/26 | Higher Order ODEs, <br>Characteristic Equations, & <br>Modeling Physical Processes | TBA | Assigned: Homework 7 | TBA |
|  | Th 3/28 | Analyzing Linear and Non-Linear 2nd-Order ODEs | TBA | - | TBA |
| **12** | Tu 4/2 | Undetermined Coefficients for Solving Linear 2nd-Order ODEs | TBA | Assigned: Homework 8 | TBA |
|  | Th 4/4 | Variation of Parameters for Solving Linear 2nd-Order ODEs | TBA | - | TBA |
| **13** | Tu 4/9 | Laplace Transforms for Solving Linear 2nd-Order ODEs | TBA | Assigned: Homework 9 | TBA |
|  | Th 4/11 | Power and Fourier Series Methods for Solving Linear 2nd-Order ODEs | TBA | - | TBA |
| **14** | Tu 4/16 | _Review_ | Exam 3 Examples | - | Exam 3 Guide |
|  | Th 4/18 | **Exam 3** | - | - | - |
| **15** | Tu 4/23 | Analyzing the Lorenz System | TBA | Assigned: Homework 10 (Optional) | TBA |
|  | Th 4/25 | Analyzing the Double Pendulum | TBA | - | TBA |
| **16** | Tu 4/30 | Final Exam Section B | - | - | - |
|  | Th 5/2 | Final Exam Section C | - | - | - |

<br>

Along with textbooks [H] and [T], and websites [C] and [P], most of the course materials (contents of worksheets and homework) of each topic was taken from these following sources:

-   Inquiry oriented differential equations (IODE) by @iode2018
-   Differential equations and linear algebra by @strang2014differential
-   Applied differential equations: The primary course by @dobrushkin2022applied
-   Differential Equations by @differentialequationsmitopencourseware

<br>

## **Books & Online Resources Lists**

```{r echo=FALSE}
refs <- bibtex::read.bib("references.bib")
```

*Click on the link to access the resources.*

### Textbooks

[H] `r capture.output(print(refs["hallstrom2018"]))` [[pdf]](https://drive.google.com/file/d/12G9cs0RQPqiXb7vjJXVH58ePJ8DZBgC1/view?usp=sharing){target="_blank"}

[T] `r capture.output(print(refs["trench2013elementary"]))` [[pdf]](https://digitalcommons.trinity.edu/cgi/viewcontent.cgi?article=1007&context=mono){target="_blank"}

### Websites

[C] `r capture.output(print(refs["clontz2022"]))`

[P] `r capture.output(print(refs["paulodenotes"]))`

<br>

## **References**

::: {#refs}
:::

<br>
